---
title: "Developer Guide"
author: "Michael E. Cotterell"
output: rmarkdown::html_vignette
description: >
  A comprehensive resource for contributing to TDCM.
vignette: >
  %\VignetteIndexEntry{Developer Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This guide is a comprehensive resource for contributing to the _R_ package **TDCM** and is intended 
for both new and experienced contributors. Contributions to TDCM are always welcome; however, to 
keep things consistent and easier to maintain, the package maintainer asks that all contributors 
follow the policies and guidelines included in this document when preparing and submitting their 
contributions to the package.

```{r, include = FALSE}
# determine the current package maintainer
tdcm_pkg <- utils::packageDescription("TDCM")
tdcm_maintainer <- stringr::str_match(tdcm_pkg$Maintainer, "(?<name>.*?) <(?<email>.*?)>")[1, ]
tdcm_maintainer_name <- tdcm_maintainer["name"]
tdcm_maintainer_email <- tdcm_maintainer["email"]
```

## Types of Contributions

TODO write.

### Issue Reporting

TODO write.

### Feature Request

TODO write.

### Code Contributions

TODO write.

## Git & Pull Requests

TODO write.

### What is Git?

TODO write.

### What is a Pull Request?

TODO write.

### Using RStudio to Prepare a Pull Request

RStudio does not have a built-in method of creating a pull request. However, it has a built-in terminal that can be used to push changes to a forked repository with a GitHub remote origin. GitHub must then be used to create a pull request.
1. On GitHub: Click the “Fork” button on the project to make changes to.
2. On RStudio: Create a new project and use Version Control with Git. Copy and paste the repository URL of your fork of the project.
3. Add the original repository as an upstream version. This should be done in RStudio’s built-in terminal. Type in: git remote add upstream https://github.com/cotterell/tdcm.git
4. Create and checkout a branch in which you will make changes.
5. Commit your changes using the “Git” tab on RStudio. Ensure that you are on the branch with your changes.
6. Push to your forked repository in the terminal: git push origin <branch name>
7. On GitHub: Click the green icon in the top left of your repository page and create a pull request with a message describing your changes.

### Using a Terminal to Prepare a Pull Request

Git supports the creation of a pull request using the “git-request-pull” command.
1. Once you have forked the original project and cloned it,  add the original repository as an upstream version:  git remote add upstream https://github.com/cotterell/tdcm.git
2. Commit your changes on a separate branch and push to the forked repository.
3. Create a pull request: git request-pull [-p] <start> <URL> [<end>]
  -p: Include patch text in your request.
  <start>: The commit to start at. This will be a commit in the upstream (original) version.
  <URL>: The repository URL to pull from; your forked repository.
  <end>: The commit to end at. This will be the final commit in your forked repository from which changes will be drawn.

## Policies

TODO write. For example, do not modify the `DESCRIPTION` file unless you are the package maintainer.

## Guidelines

TODO write.

## Maintainer Notes

This section includes notes that are intended for the package maintainer only. At the time of this
writing, the package maintainer is `r tdcm_maintainer_name` based on the package's `DESCRIPTION` 
file.

### GitHub Actions

#### R-CMD-check

[![R-CMD-check](https://github.com/cotterell/tdcm/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/cotterell/tdcm/actions/workflows/R-CMD-check.yaml)

This workflow runs `R-CMD-check` using the three major operating systems (linux, macOS and Windows) 
with the current, development, and previous versions of R.

#### pkgdown

[![pkgdown](https://github.com/cotterell/tdcm/actions/workflows/pkgdown.yaml/badge.svg)](https://github.com/cotterell/tdcm/actions/workflows/pkgdown.yaml)

This workflow builds the package website using `pkgdown::build_site_github_pages()`, then pushes the 
built package website to GitHub Pages. The inclusion of
[`workflow_dispatch`](https://docs.github.com/en/actions/learn-github-actions/events-that-trigger-workflows#workflow_dispatch)
means that the workflow can 
[be run manually](https://docs.github.com/en/actions/managing-workflow-runs/manually-running-a-workflow)
or [triggered via the GitHub REST API](https://docs.github.com/en/rest/reference/actions/#create-a-workflow-dispatch-event).

- Release website: [https://cotterell.github.io/tdcm/](https://cotterell.github.io/tdcm/)
- Development website: [https://cotterell.github.io/tdcm/dev/](https://cotterell.github.io/tdcm/dev/)
  
#### pr-commands

[![pr-commands](https://github.com/cotterell/tdcm/actions/workflows/pr-commands.yaml/badge.svg)](https://github.com/cotterell/tdcm/actions/workflows/pr-commands.yaml) 

This workflow adds `/document` and `/style` commands for pull requests.

#### test-coverage

[![test-coverage](https://github.com/cotterell/tdcm/actions/workflows/test-coverage.yaml/badge.svg)](https://github.com/cotterell/tdcm/actions/workflows/test-coverage.yaml)

This workflow uses `covr::codecov()` to query the test coverage of the package and upload the result 
to [https://app.codecov.io/gh/cotterell/tdcm](https://app.codecov.io/gh/cotterell/tdcm).
 
### Prepare a Release

TODO write.

## References

Chacon, S., & Straub, B. (November 2014). 
_Pro Git_. 
Second Edition. Apress. ISBN: 978-1484200773.
[https://git-scm.com/book/en/v2](https://git-scm.com/book/en/v2)

Wickham, H., & Bryan, J. (July 2023). 
_R Packages: Organize, Test, Document, and Share Your Code._ 
Second Edition. O'Reilly Media. ISBN: 9781098134945.
[https://r-pkgs.org/](https://r-pkgs.org/)

Wickham, H., Hester, J., Chang. W., & Bryan, J. (2022). 
_devtools: Tools to Make Developing R Packages Easier_. 
[https://devtools.r-lib.org/](https://devtools.r-lib.org/), 
[https://github.com/r-lib/devtools](https://github.com/r-lib/devtools)
