---
title: "Developer Guide"
author: "Michael E. Cotterell"
output: rmarkdown::html_vignette
description: >
  A comprehensive resource for contributing to TDCM.
vignette: >
  %\VignetteIndexEntry{Developer Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This guide is a comprehensive resource for contributing to the _R_ package **TDCM** and is intended 
for both new and experienced contributors. Contributions to TDCM are always welcome; however, to 
keep things consistent and easier to maintain, the package maintainer asks that all contributors 
follow the policies and guidelines included in this document when preparing and submitting their 
contributions to the package.

```{r, include = FALSE}
# determine the current package maintainer
tdcm_pkg <- utils::packageDescription("TDCM")
tdcm_maintainer <- stringr::str_match(tdcm_pkg$Maintainer, "(?<name>.*?) <(?<email>.*?)>")[1, ]
tdcm_maintainer_name <- tdcm_maintainer["name"]
tdcm_maintainer_email <- tdcm_maintainer["email"]
```

## Types of Contributions

TODO write.

### Issue Reporting

TODO write.

### Feature Request

TODO write.

### Code Contributions

For your contribution to be accepted, it needs to follow a consistent style, be well documented, 
and include unit tests. You can ensure your code meets our standards by following the steps 
outlined below. All package instructions are written for use in RStudio.

To install all necessary packages, run the following command:
```{r, include = TRUE}
install.packages(c("styler", "roxygen2", "testthat", "covr"))
```
Packages: [styler](https://cran.r-project.org/web/packages/styler/index.html), 
[roxygen2](https://cran.r-project.org/web/packages/roxygen2/index.html), 
[testthat](https://cran.r-project.org/web/packages/testthat/index.html), 
[covr](https://cran.r-project.org/web/packages/covr/index.html)

#### 1. Code Style

Code style is how your code looks and it's an important part of contributing to open source 
projects. Using a consistent style makes your code easier for yourself and future collaborators to 
understand. Follow the steps below for each file in your pull request.

1. Open a file to style.
2. Press Ctrl + Shift + P to open the command palette.
3. Search for "style actvive file" and click on it.

#### 2. Documentation

Accurate documentation is essential for maintaining an open source project. It allows developers to 
quickly understand the purpose, functionality, and usage of different parts of the codebase, making 
it easier for new collaborators to contribute and saving developers time when reviewing pull 
requests. Follow the steps below for each function you write.

1. Select Build > Configure Build Tools.
2. Check the box for "Generate documentation with Roxygen" and click OK (on both windows).
3. Put your cursor inside the body of a function you want to generate documentation for.
4. Select Code > Insert Roxygen Skeleton.
5. Edit the Roxygen comment with a description of your code.

#### 3. Unit Tests

A unit test is a function that validates the expected behavior of another function. The goal is to 
make sure that each part of your code works independently before merging it with a larger codebase.

1. Write a function. Here's a simple example:
```{r, include = TRUE}
#' Multiplies two numbers.
#'
#' @param y number to be multiplied
#' @param x number to be multiplied
#'
#' @return result the result
multiply <- function(x, y) {
  result <- x * y
  return(result)
}
```

2. Write a unit test that varifies the accuracy of your function. A list of function expectations 
can be found [here](https://testthat.r-lib.org/reference/). Create a script for your tests and put 
it in the ```test/testthat``` directory.

```{r, include = TRUE}
#' Usage
test_that(desc, code)

#' Example
test_that("multiply works", {
  expect_equal(multiply(2, 4), 8)
})
```

3. Code coverage is the amount of code that is checked by your unit tests. A minimum coverage of 
80% is required for contributions to TDCM. Run the following commands to view a list of all R 
scripts in the package and their coverage, as well as the total coverage for the entire package. 
Make sure all the scripts you wrote are above 80% before submitting a pull request.

```{r, include = TRUE}
> library(covr)
> package_coverage()
```

## Git & Pull Requests

TODO write.

### What is Git?

TODO write.

### What is a Pull Request?

TODO write.

### Using RStudio to Prepare a Pull Request

RStudio does not have a built-in method of creating a pull request. However, it has a built-in terminal that can be used to push changes to a forked repository with a GitHub remote origin. GitHub must then be used to create a pull request.
1. On GitHub: Click the “Fork” button on the project to make changes to.
2. On RStudio: Create a new project and use Version Control with Git. Copy and paste the repository URL of your fork of the project.
3. Add the original repository as an upstream version. This should be done in RStudio’s built-in terminal. Type in: git remote add upstream https://github.com/cotterell/tdcm.git
4. Create and checkout a branch in which you will make changes.
5. Commit your changes using the “Git” tab on RStudio. Ensure that you are on the branch with your changes.
6. Push to your forked repository in the terminal: git push origin <branch name>
7. On GitHub: Click the green icon in the top left of your repository page and create a pull request with a message describing your changes.

### Using a Terminal to Prepare a Pull Request

Git supports the creation of a pull request using the “git-request-pull” command.
1. Once you have forked the original project and cloned it,  add the original repository as an upstream version:  git remote add upstream https://github.com/cotterell/tdcm.git
2. Commit your changes on a separate branch and push to the forked repository.
3. Create a pull request: git request-pull [-p] <start> <URL> [<end>]
  -p: Include patch text in your request.
  <start>: The commit to start at. This will be a commit in the upstream (original) version.
  <URL>: The repository URL to pull from; your forked repository.
  <end>: The commit to end at. This will be the final commit in your forked repository from which changes will be drawn.

## Policies

TODO write. For example, do not modify the `DESCRIPTION` file unless you are the package maintainer.

## Guidelines

TODO write.

## Maintainer Notes

This section includes notes that are intended for the package maintainer only. At the time of this
writing, the package maintainer is `r tdcm_maintainer_name` based on the package's `DESCRIPTION` 
file.

### GitHub Actions

#### R-CMD-check

[![R-CMD-check](https://github.com/cotterell/tdcm/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/cotterell/tdcm/actions/workflows/R-CMD-check.yaml)

This workflow runs `R-CMD-check` using the three major operating systems (linux, macOS and Windows) 
with the current, development, and previous versions of R.

#### pkgdown

[![pkgdown](https://github.com/cotterell/tdcm/actions/workflows/pkgdown.yaml/badge.svg)](https://github.com/cotterell/tdcm/actions/workflows/pkgdown.yaml)

This workflow builds the package website using `pkgdown::build_site_github_pages()`, then pushes the 
built package website to GitHub Pages. The inclusion of
[`workflow_dispatch`](https://docs.github.com/en/actions/learn-github-actions/events-that-trigger-workflows#workflow_dispatch)
means that the workflow can 
[be run manually](https://docs.github.com/en/actions/managing-workflow-runs/manually-running-a-workflow)
or [triggered via the GitHub REST API](https://docs.github.com/en/rest/reference/actions/#create-a-workflow-dispatch-event).

- Release website: [https://cotterell.github.io/tdcm/](https://cotterell.github.io/tdcm/)
- Development website: [https://cotterell.github.io/tdcm/dev/](https://cotterell.github.io/tdcm/dev/)
  
#### pr-commands

[![pr-commands](https://github.com/cotterell/tdcm/actions/workflows/pr-commands.yaml/badge.svg)](https://github.com/cotterell/tdcm/actions/workflows/pr-commands.yaml) 

This workflow adds `/document` and `/style` commands for pull requests.

#### test-coverage

[![test-coverage](https://github.com/cotterell/tdcm/actions/workflows/test-coverage.yaml/badge.svg)](https://github.com/cotterell/tdcm/actions/workflows/test-coverage.yaml)

This workflow uses `covr::codecov()` to query the test coverage of the package and upload the result 
to [https://app.codecov.io/gh/cotterell/tdcm](https://app.codecov.io/gh/cotterell/tdcm).
 
### Prepare a Release

TODO write.

## References

Chacon, S., & Straub, B. (November 2014). 
_Pro Git_. 
Second Edition. Apress. ISBN: 978-1484200773.
[https://git-scm.com/book/en/v2](https://git-scm.com/book/en/v2)

Wickham, H., & Bryan, J. (July 2023). 
_R Packages: Organize, Test, Document, and Share Your Code._ 
Second Edition. O'Reilly Media. ISBN: 9781098134945.
[https://r-pkgs.org/](https://r-pkgs.org/)

Wickham, H., Hester, J., Chang. W., & Bryan, J. (2022). 
_devtools: Tools to Make Developing R Packages Easier_. 
[https://devtools.r-lib.org/](https://devtools.r-lib.org/), 
[https://github.com/r-lib/devtools](https://github.com/r-lib/devtools)
