---
title: "Developer Guide"
author: "Michael E. Cotterell"
output: rmarkdown::html_vignette
description: >
  A comprehensive resource for contributing to TDCM.
vignette: >
  %\VignetteIndexEntry{Developer Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This guide is a comprehensive resource for contributing to the _R_ package **TDCM** and is intended 
for both new and experienced contributors. Contributions to TDCM are always welcome; however, to 
keep things consistent and easier to maintain, the package maintainer asks that all contributors 
follow the policies and guidelines included in this document when preparing and submitting their 
contributions to the package.

```{r, include = FALSE}
# determine the current package maintainer
tdcm_pkg <- utils::packageDescription("TDCM")
tdcm_maintainer <- stringr::str_match(tdcm_pkg$Maintainer, "(?<name>.*?) <(?<email>.*?)>")[1, ]
tdcm_maintainer_name <- tdcm_maintainer["name"]
tdcm_maintainer_email <- tdcm_maintainer["email"]
```

## Types of Contributions

Would you like to help contribute to TDCM? We're more than happy to work with our community to improve our user experience!
There are various ways to contribute: 

* Issue Reporting - report any bugs or issues you come across so we can fix them!
* Feature Requests - let us know about any features you'd like to see in TDCM!
* Code Contributions - help us work on TDCM by fixing bugs or adding features yourself!

You can learn more about each of these types of contributions below!

### Issue Reporting

If you notice any errors, bugs, or unexpected behaviors in the package, please use the Issues feature on GitHub! This will allow us to respond promptly and keep all required fixes in one place. To do so, log in to your GitHub account (or create one if you do not already have one) and follow the instructions below:

1. Navigate to the [TDCM GitHub Issues page](https://github.com/cotterell/tdcm/issues)
2. Use the search box to see if the issue you encountered has already been reported.
    * If the issue has been reported, leave a reply with as much detail as possible explaining how you encountered the issue. Try to be as specific as possible so the issue can be easily reproduced!
    * If the issue has not been reported yet, make your own!:
        1. Click the green “New Issue” button
        2. Click “Get Started” next to “Bug Report” [TODO: make this link once templates are added to main repository]
        3. Click “Submit New Issue” once you’ve finished filling out the form!
3. Once you’ve reported the issue, developers may reply to your report asking for more details. Please provide as much information as possible to help ensure the issue is resolved quickly!

### Feature Request

Have any features you’d like us to add? Please let us know using the Issues feature on GitHub!

> Issues is a feature on GitHub that streamlines communication between users and developers and consolidates all project tasks in one place. While this of course includes bugs, it is often used for managing feature requests and improvements, allowing it to serve as a "To-Do List" developers can use to know what they should be focusing on.

To create an Issue, log in to your GitHub account (or create one if you do not already have one) and follow the instructions below:

1. Navigate to the [TDCM GitHub Issues page](https://github.com/cotterell/tdcm/issues)
2. Use the search box to see if the feature you’re thinking of has already been requested.
	* If there is already a request, please interact with it!
		* You can use the emoji icon to add a reaction, such as a “thumbs up” to show your support! This will let us know which features to prioritize by showing how many people want a particular feature. 
		* Additionally, feel free to voice your opinions in the replies of the request page! You can add additional details you’d like to see implemented, or simply express your desire for the feature to be added!
	* If there is no existing request for the feature, you can make your own:
		1. Click the green “New Issue” button
		2. Click “Get Started” next to “Feature Request” [TODO: make this a link once templates are added to main repository]
		3. Click “Submit New Issue” once you’ve finished filling out the form!
3. Once you’ve posted your feature request, developers may reply to your request asking for more details. Please provide as much information as possible so the developers know what you’re looking for!

### Code Contributions

For your contribution to be accepted, it needs to follow a consistent style, be well documented, 
and include unit tests. You can ensure your code meets our standards by following the steps 
outlined below. All package instructions are written for use in RStudio.

To install all necessary packages, run the following command:

Packages: [styler](https://cran.r-project.org/web/packages/styler/index.html), 
[roxygen2](https://cran.r-project.org/web/packages/roxygen2/index.html), 
[testthat](https://cran.r-project.org/web/packages/testthat/index.html), 
[covr](https://cran.r-project.org/web/packages/covr/index.html)

#### 1. Code Style

Code style is how your code looks and it's an important part of contributing to open source 
projects. Using a consistent style makes your code easier for yourself and future collaborators to 
understand. Follow the steps below for each file in your pull request.

1. Open a file to style.
2. Press Ctrl + Shift + P to open the command palette.
3. Search for "style actvive file" and click on it.

#### 2. Documentation

Accurate documentation is essential for maintaining an open source project. It allows developers to 
quickly understand the purpose, functionality, and usage of different parts of the codebase, making 
it easier for new collaborators to contribute and saving developers time when reviewing pull 
requests. Follow the steps below for each function you write.

1. Select Build > Configure Build Tools.
2. Check the box for "Generate documentation with Roxygen" and click OK (on both windows).
3. Put your cursor inside the body of a function you want to generate documentation for.
4. Select Code > Insert Roxygen Skeleton.
5. Edit the Roxygen comment with a description of your code.

#### 3. Unit Tests

A unit test is a function that validates the expected behavior of another function. The goal is to 
make sure that each part of your code works independently before merging it with a larger codebase.

1. Write a function. Here's a simple example:


2. Write a unit test that varifies the accuracy of your function. A list of function expectations 
can be found [here](https://testthat.r-lib.org/reference/). Create a script for your tests and put 
it in the ```test/testthat``` directory.



3. Code coverage is the amount of code that is checked by your unit tests. A minimum coverage of 
80% is required for contributions to TDCM. Run the following commands to view a list of all R 
scripts in the package and their coverage, as well as the total coverage for the entire package. 
Make sure all the scripts you wrote are above 80% before submitting a pull request.


## Git & Pull Requests

The below section describes what Git is and how to use it to help contribute to this repository.

### What is Git?

[Git](https://git-scm.com/book/en/v2/Getting-Started-What-is-Git%3F) is a version control system that allows developers to track changes in their code over time. It's like a time machine for your project files, enabling you to save different versions (called "commits") of your code. Each commit records what changes were made, by whom, and when, making it easy to revert back to older versions if something goes wrong with the current code.

This is a breakdown of key features and concepts in Git:
[Repository](https://git-scm.com/book/en/v2/Git-Basics-Getting-a-Git-Repository) (Repo): This is where all your project’s files and their history are stored. It can be located on your computer (local repository) or on a remote server (remote repository).
[Commit](https://git-scm.com/book/en/v2/Git-Basics-Recording-Changes-to-the-Repository): A commit is like a snapshot of your project at a specific point in time. When you make changes to your files, you 'commit' these changes, which saves a record of what was altered, allowing you to track the evolution of your project.
[Branch](https://git-scm.com/book/en/v2/Git-Basics-Recording-Changes-to-the-Repository): Branches allow you to create separate versions of your project within the same repository. You might use a branch to develop a new feature while keeping the main project (usually called the "master" or "main" branch) stable.
[Merge](https://git-scm.com/book/en/v2/Git-Basics-Recording-Changes-to-the-Repository): When you've finished working in a branch, you can merge your changes back into the main branch. This combines the histories of both branches, integrating the new features or changes into the main project.
[Pull](https://git-scm.com/book/en/v2/GitHub-Maintaining-a-Project): Pulling refers to downloading changes from a remote repository to your local one. This is useful when working in teams, as it lets you see and integrate the changes made by others.
[Push](https://git-scm.com/docs/git-push): Pushing refers to uploading your local repository changes to a remote repository. This is how you share your work with others or back up your current state to a remote location.
Using [Git](https://git-scm.com/video/what-is-git), [multiple](https://git-scm.com/video/what-is-version-control) developers can work on the same project simultaneously without interfering with each other’s work, as they can work in different branches and merge changes only when ready. This collaborative feature, along with the ability to track every change, makes Git an essential tool for modern software development.

What is the difference between Git and Github?

[Git](https://git-scm.com/):
Definition: Git is a version control system (VCS) that helps you track changes in your code files. It allows multiple developers to work together on the same project effectively and efficiently by providing tools to manage differing versions of code.
Functionality: Git is used to create branches, handle merges, revert to previous states, and more, directly from your local machine or a server without needing an internet connection.
Local vs. Remote: Git operates primarily on your local system, managing your local repository. However, it can synchronize with remote repositories.
[GitHub](https://github.com/):
Definition: GitHub is a hosting service for Git repositories. It's a platform where you can upload your Git repositories and collaborate with others.
Functionality: Beyond the basic functionalities of Git, GitHub provides additional features such as bug tracking, feature requests, task management, and wikis for project documentation.
Collaboration: GitHub is designed to facilitate easier collaboration among team members on a project. It offers graphical tools for managing pull requests, code reviews, and even project management tools like issues and project boards.
In essence, Git is the tool that enables version control, while GitHub is a service that hosts your repositories and adds extra features to enhance collaboration among team members. Both work hand-in-hand to support the development process, but they serve different roles within that framework.




### What is a Pull Request?

A [Pull](https://docs.github.com/en/pull-requests/collaborating-with-pull-requests/reviewing-changes-in-pull-requests/about-pull-request-reviews) Request (PR), also known as a [Merge](https://git-scm.com/book/en/v2/Git-Branching-Basic-Branching-and-Merging) Request in some systems like GitLab, is a feature used in Git repositories that lets developers notify team members about changes they've pushed to a branch in a repository on platforms like GitHub, Bitbucket, or GitLab. Essentially, it's a request to review and then merge this branch into another branch, typically the main project branch.
Here's how it works in simple terms:
[Branch Creation](https://docs.github.com/en/pull-requests/collaborating-with-pull-requests/proposing-changes-to-your-work-with-pull-requests/creating-a-pull-request): First, a developer creates a branch from the main codebase to work on a new feature, a bug fix, or any updates. This separation ensures that the main codebase remains unaffected until the new changes are reviewed and approved.
[Making Changes and Committing](https://git-scm.com/book/en/v2/Git-Basics-Recording-Changes-to-the-Repository): The developer makes necessary changes in this new branch, commits these changes, and pushes the branch to a remote repository.
Opening a Pull Request: Once the branch is updated on the remote repository, the developer opens a pull request. This PR is essentially a request to the rest of the team to review the changes made in the branch before it gets merged into the main branch.
Review and Discussion: Team members review the changes, discuss potential improvements, and may request further [changes](https://docs.github.com/en/pull-requests/collaborating-with-pull-requests/proposing-changes-to-your-work-with-pull-requests/about-comparing-branches-in-pull-requests) if needed. This process is crucial for maintaining code quality and ensuring that the integration aligns with the project's overall direction.
Approval and [Merge](https://docs.github.com/en/pull-requests/collaborating-with-pull-requests/addressing-merge-conflicts/resolving-a-merge-conflict-on-github): After the changes are reviewed and approved by the necessary team members or according to the project's guidelines, the pull request can be merged. This means that the changes in the branch will be integrated into the target branch, usually the main or master branch.
Closure and Cleanup: Once merged, the pull request is closed, and it's common to delete the branch to keep the repository clean and manageable.
Pull Requests are a vital part of the collaborative development process, enabling transparency, improving code quality, and helping manage contributions to a project efficiently. They facilitate a peer review system which is beneficial for both team cohesion and project integrity.



### Using RStudio to Prepare a Pull Request

RStudio does not have a built-in method of creating a pull request. However, it has a built-in terminal that can be used to push changes to a forked repository with a GitHub remote origin. GitHub must then be used to create a pull request.
1. On GitHub: Click the “Fork” button on the project to make changes to.
2. On RStudio: Create a new project and use Version Control with Git. Copy and paste the repository URL of your fork of the project.
3. Add the original repository as an upstream version. This should be done in RStudio’s built-in terminal. Type in: git remote add upstream https://github.com/cotterell/tdcm.git
4. Create and checkout a branch in which you will make changes.
5. Commit your changes using the “Git” tab on RStudio. Ensure that you are on the branch with your changes.
6. Push to your forked repository in the terminal: git push origin <branch name>
7. On GitHub: Click the green icon in the top left of your repository page and create a pull request with a message describing your changes.

### Using a Terminal to Prepare a Pull Request

You can use [GitHub’s CLI (downloadable)](https://cli.github.com) to fork a repository and create a pull request within it.
Fork the original project:[[gh repo fork [<repository>](https://cli.github.com/manual/gh_repo_fork) 
Commit your changes on a separate branch.
Create a pull request: [gh pr create](https://cli.github.com/manual/gh_pr_create) 

#### Continuous Integration and Pull Requests
* This TDCM expansion utilizes continuous integration via GitHub’s **“Actions” column**. This means that a script within the project will automatically run various **checks** on a pull request and notify you of any problems on your **GitHub profile**. Once these are resolved, provided that they do not require further inspection, your pull request will be automatically integrated.  

## Policies

TODO write. For example, do not modify the `DESCRIPTION` file unless you are the package maintainer.

## Guidelines

General Guidelines


Adherence to Standards:

- Follow the coding style and documentation standards outlined in this guide.

- Ensure your contributions align with the project’s goals and design principles.


Quality Over Quantity:

- Focus on delivering high-quality contributions that are well-tested and well-documented.

- Avoid making changes that are unnecessary or do not add significant value to the project.

Respect and Collaboration:

- Communicate respectfully and professionally with other contributors and maintainers.

- Be open to feedback and willing to make revisions based on reviews.

Code Contributions
Consistency:

- Use consistent code style throughout your contributions. 
Adhere to the conventions used in the existing codebase.

- Use the styler package to format your code.

Documentation:

- Document your code using roxygen2 comments. 

- This helps other developers understand the purpose and usage of your functions.

- Update any relevant documentation files if your changes affect the usage or behavior of the package.

Testing:

- Write unit tests for your code using the testthat package to ensure it works as expected.

- Aim for a minimum of 80% code coverage for your contributions. Use the covr package to check coverage.

Submission:

- Make sure your code passes all checks and tests before submitting a pull request.

- Provide a clear and detailed description of the changes in your pull request.

Issue Reporting
Clarity and Detail:

- Provide a clear and concise description of the issue.

- Include details such as steps to reproduce the issue, expected behavior, and any relevant screenshots or error messages.
Existing Issues:

- Before creating a new issue, check the existing issues to see if the problem has already been reported.

- If the issue already exists, add any additional information you have and indicate your interest in a resolution.

Feature Requests
Feasibility and Impact:

- Ensure that your feature request is feasible and aligns with the project’s goals.

- Clearly explain the benefits and potential impact of the proposed feature.


Community Involvement:

- Engage with the community by participating in discussions about the feature.

- Be responsive to feedback and willing to refine your proposal based on community input.

Git & Pull Requests

Branch Management:

- Create a new branch for each feature or bug fix to keep your work organized and separate from the main codebase.

- Use descriptive names for your branches to indicate their purpose.

Commit Practices:

- Write clear and concise commit messages that describe the changes made.

- Make small, incremental commits rather than large, monolithic ones.

Pull Request Submission:

- Ensure your branch is up-to-date with the main branch before opening a pull request.

- Provide a detailed description of the changes, including the problem being addressed and the solution implemented.

- Be responsive to feedback from reviewers and make necessary revisions promptly.

Continuous Integration
Automated Checks:

- Ensure your code passes all automated checks configured in the GitHub Actions workflows.

- Address any issues reported by the CI tools before requesting a review.

Maintaining Build Health:

- Avoid breaking changes that could disrupt the build or affect other contributors.

- Regularly sync your branch with the main branch to incorporate the latest changes and resolve conflicts.

By following these guidelines, you contribute to maintaining the quality and integrity of the TDCM project, ensuring a positive experience for all contributors and users.


## Maintainer Notes

This section includes notes that are intended for the package maintainer only. At the time of this
writing, the package maintainer is `r tdcm_maintainer_name` based on the package's `DESCRIPTION` 
file.

### GitHub Actions

#### R-CMD-check

[![R-CMD-check](https://github.com/cotterell/tdcm/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/cotterell/tdcm/actions/workflows/R-CMD-check.yaml)

This workflow runs `R-CMD-check` using the three major operating systems (linux, macOS and Windows) 
with the current, development, and previous versions of R.

#### pkgdown

[![pkgdown](https://github.com/cotterell/tdcm/actions/workflows/pkgdown.yaml/badge.svg)](https://github.com/cotterell/tdcm/actions/workflows/pkgdown.yaml)

This workflow builds the package website using `pkgdown::build_site_github_pages()`, then pushes the 
built package website to GitHub Pages. The inclusion of
[`workflow_dispatch`](https://docs.github.com/en/actions/learn-github-actions/events-that-trigger-workflows#workflow_dispatch)
means that the workflow can 
[be run manually](https://docs.github.com/en/actions/managing-workflow-runs/manually-running-a-workflow)
or [triggered via the GitHub REST API](https://docs.github.com/en/rest/reference/actions/#create-a-workflow-dispatch-event).

- Release website: [https://cotterell.github.io/tdcm/](https://cotterell.github.io/tdcm/)
- Development website: [https://cotterell.github.io/tdcm/dev/](https://cotterell.github.io/tdcm/dev/)
  
#### pr-commands

[![pr-commands](https://github.com/cotterell/tdcm/actions/workflows/pr-commands.yaml/badge.svg)](https://github.com/cotterell/tdcm/actions/workflows/pr-commands.yaml) 

This workflow adds `/document` and `/style` commands for pull requests.

#### test-coverage

[![test-coverage](https://github.com/cotterell/tdcm/actions/workflows/test-coverage.yaml/badge.svg)](https://github.com/cotterell/tdcm/actions/workflows/test-coverage.yaml)

This workflow uses `covr::codecov()` to query the test coverage of the package and upload the result 
to [https://app.codecov.io/gh/cotterell/tdcm](https://app.codecov.io/gh/cotterell/tdcm).
 
### Prepare a Release

TODO write.

## References

Chacon, S., & Straub, B. (November 2014). 
_Pro Git_. 
Second Edition. Apress. ISBN: 978-1484200773.
[https://git-scm.com/book/en/v2](https://git-scm.com/book/en/v2)

Wickham, H., & Bryan, J. (July 2023). 
_R Packages: Organize, Test, Document, and Share Your Code._ 
Second Edition. O'Reilly Media. ISBN: 9781098134945.
[https://r-pkgs.org/](https://r-pkgs.org/)

Wickham, H., Hester, J., Chang. W., & Bryan, J. (2022). 
_devtools: Tools to Make Developing R Packages Easier_. 
[https://devtools.r-lib.org/](https://devtools.r-lib.org/), 
[https://github.com/r-lib/devtools](https://github.com/r-lib/devtools)
